Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-11-12 paper/main.tex
spacing

The rules {\sc Q:IncP}\iffull{, {\sc Q:DecP},}{} and {\sc Q:Inc}
describe how the potential is distributed after a size change of
a variable.
\iffull{The rule {\sc Q:IncP} is for increments $x \gets x + y$
and {\sc Q:DecP} is for decrements $x \gets x - y$, they both apply only
when we can deduce from the logical context $\Gamma$ that $y \ge 0$.
Of course, we have symmetrical rules {\sc Q:IncN} and {\sc Q:DecN}
that can be applied if $y$ is not positive.}{The rule {\sc Q:IncP} is for increments
$x \gets x+y$, it applies when we can deduce from the logical context
$\Gamma$ that $y \ge 0$.  Of course, we have symmetrical rules that
can be applied for decrements and when $y$ is not positive.}
The rules are all equivalent in the case where $y=0$.
The rule {\sc Q:Inc} can be applied
if we cannot deduce any information about $y$.  In the implementation,
the rules for increments and decrements are combined into one syntax
directed rule\iffull{ where the specialized rules take priority over {\sc Q:Inc}.}{.}

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x - 10 \geq 0)~\{x=x{-}10; tick(5);\}~else~break;}}}
\begin{figure*}
\vspace{-.6cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (x {<} 10, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (\cdot, P^\text{de})
          }
        }
        { (x {<} 10, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (x {<} 10, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (x {<} 10, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (x {<} 10, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (x {<} 10, B^\text{br}); (x {<} 10,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (x {<} 10, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (x {<} 10, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (x {<} 10, P)
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(\cdot)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced by the automatic tool.}
\label{fig:derivation}
\end{figure*}

@@

\begin{figure}
\center
\begin{minipage}[b]{0.8\linewidth}
\begin{lstlisting}[numbers=left]
${\color{blue}%
  \mbox{logical state invariant } \{ na = \#_1(a)} \}$
while (k > 0) {
  x=0;
  while (x < N && a[x] == 1) {
    @assert(na > 0);@
    a[x]=0; @na--;@
    tick(1); x++; }
  if (x < N) { a[x]=1; @na++;@ tick(1); }
  k--;
}
\end{lstlisting}
\end{minipage}
\caption{Assisted bound derivation using logical state.  We write
  $\#_1(a)$ for $\# \{ i \mid 0 {\le} i {<} N {\land} a[i] {=} 1 \}$ and use the
  tick metric.  The derived bound  is
  $2|[0,k]| + |[0,na]|$.
  }
\label{fig:xmplincaux}
\end{figure}

>>>>

The rules {\sc Q:IncP}\iffull{, {\sc Q:DecP},}{} and {\sc Q:Inc}
describe how the potential is distributed after a size change of
a variable.
\iffull{The rule {\sc Q:IncP} is for increments $x \gets x + y$
and {\sc Q:DecP} is for decrements $x \gets x - y$, they both apply only
when we can deduce from the logical context $\Gamma$ that $y \ge 0$.
Of course, we have symmetrical rules {\sc Q:IncN} and {\sc Q:DecN}
that can be applied if $y$ is not positive.}{The rule {\sc Q:IncP} is for increments
$x \gets x+y$, it applies when we can deduce from the logical context
$\Gamma$ that $y \ge 0$.  Of course, we have symmetrical rules that
can be applied for decrements and when $y$ is not positive.}
The rules are all equivalent in the case where $y=0$.
The rule {\sc Q:Inc} can be applied
if we cannot deduce any information about $y$.  In $\toolname$,
the rules for increments and decrements are combined into one rule\iffull{ where the specialized rules take priority over {\sc Q:Inc}.}{.}

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x - 10 \geq 0)~\{x=x{-}10; tick(5);\}~else~break;}}}
\begin{figure*}[t]
\vspace{-.6cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (x {<} 10, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (\cdot, P^\text{de})
          }
        }
        { (x {<} 10, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (x {<} 10, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (x {<} 10, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (x {<} 10, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (x {<} 10, B^\text{br}); (x {<} 10,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (x {<} 10, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (x {<} 10, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (x {<} 10, P)
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(\cdot)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced by the automatic tool.}
\label{fig:derivation}
\end{figure*}

@@

\begin{figure}[t]
\center
\begin{minipage}[b]{0.8\linewidth}
\begin{lstlisting}[numbers=left]
${\color{blue}%
  \mbox{logical state invariant } \{ na = \#_1(a)} \}$
while (k > 0) {
  x=0;
  while (x < N && a[x] == 1) {
    @assert(na > 0);@
    a[x]=0; @na--;@
    tick(1); x++; }
  if (x < N) { a[x]=1; @na++;@ tick(1); }
  k--;
}
\end{lstlisting}
\end{minipage}
\caption{Assisted bound derivation using logical state.  We write
  $\#_1(a)$ for $\# \{ i \mid 0 {\le} i {<} N {\land} a[i] {=} 1 \}$ and use the
  tick metric.  The derived bound  is
  $2|[0,k]| + |[0,na]|$.
  }
\label{fig:xmplincaux}
\end{figure}
